### 2017-01-13 response-to-reviewers.md

> Reviewer 1
> 
> This manuscript provides rules aimed at novice programmers to help
> them write more robust bioinformatics programs. The rules are all
> sensible, and biologists turned programmers (and some people with
> formal software engineering training) would do well to read them.
> 
> With any set of guidelines, and especially guidelines about
> programming, there is room for subjectivity and differences of
> opinion, so here are mine. The authors are free to include or exclude
> these comments in their revision.

>  Robustness
>
>The rules are framed in terms of increasing robustness.
>
> The piece starts out defining robustness in terms of portability,
> specifically outside the programmer's own environment - but then goes
> on to define it more broadly in terms of multiple criteria such as
> being kept under version control (which I don't think is a necessary
> condition for robustness - many robust programs were made before VCSs
> although I think using VCSs is excellent advice for a modern
> programmer)
> 
> The definition here is somewhat different from Wikipedia's, which defines it as:
> 
> "the ability of a computer system to cope with errors during execution[1][2] and cope with erroneous input."
> 
> I am not sure of the best name for the property the authors are aiming at, but it seems some combination of robustness, maintainability, quality, predictability - essentially better software.

Thanks for the feedback. We have clarified and condensed our definition of 'robust' software in the introduction, in particular removing the reference to version control.

>     Documentation and maintainability.
> 
> I find the relegation of documentation to an additional note rather
> than a rule surprising. It's also telling the authors don't call out
> maintainability as a desirable property (except in passing in the
> conclusions). If software is not thrown away, it frequently needs to
> be updated - e.g. to accomodate changes in input file types. This
> maintenance task often falls on people who are not the original
a> uthor. If software does not have adequate documentation, it will be
> harder to maintain (either by the original author or by others).

We've promoted documentation to a full rule (rule #1 in fact).

>    Good software development practice
> 
> The rules do not mention anything about good software engineering
> practice, such as structured programming. While it's good to not be
> too specific and dictate specific fads or paradigms (and the intended
> audience is individual programmers not teams), no one would disagree
> that some kind of good structured programming practices are
> required to write programs where clarity, maintainability,
> extensibility and quality are of concern.
> 
> In general, the program should not repeat large pieces of code with
> the same logic - these should be abstracted into
> subroutines/functions/methods/classes etc. The program should not be a
> giant monolithic main() method. Behavior of subroutines should not be
> controlled through globals. etc.
> 
> While it's not necessary to replicate Knuth's three volumes, I think
> good software engineering practice deserves to be a rule, or at least
> called out.

 We've added a paragraph to the introduction to mention code quality, maintainability, unit tests, and version control as important but out of scope.

> Reproducibility
> 
> Reproducibility of software and workflows is increasingly in the
> spotlight. There are passing mentions of reproducibility, and it is
> implicit in many of the guidelines, but it would be good to call it
> out explicitly as a goal beyond making the software robust. For
> example, in the list on p2.
> 
> People like Titus Brown have written a lot about this in blogs. It
> would be useful to reference some of this.

We've added more mention of reproducibility and citations to posts written by C. Titus Brown and by Gaël Varoquaux.

> README
> 
> Excellent advice, although there is some overlap with 2.

These two rules have been combined.

> Usage info
> 
> Good advice, but I wonder if 2 and 8 could be combined?

We consider usage to be part of the documentation, but a section of this rule referred to checking input parameters and echoing out parameters to the command line. These sections have been moved to the rules on controlling common operations and producing identical results.

> Versioning
> 
> All good advice.
> 
> As the authors note, GitHub makes this very easy. The advice is a bit succinct:
> 
> "as simple as adding an appropriate commit message or tag to version control"
> 
> As this article seems to be aimed as novice programmers who may not
> have been immersed in good, modern, open software development
> practice, it would be a good idea to include more details or pointers here.
> 
> In fact, I would recommend every programmer learns how to use a VCS
> and a hosted VCS like GitHub/GitLab/Bitbucket, to commit early and
> commit often, to use issue trackers, etc. Not just a matter of making
> public releases (which is important) but also as a matter of good
> practice even for software not intended to be released.

We think it's important too, but it's a tough sell in some crowds. We've added a reference to another educational article, "A Quick Introduction to Version Control with Git and GitHub".

> Reuse software
> 
> At first I assumed this section was about reusing software
> libraries, but it transpired to be about reusing executable programs.
> 
> This seems an odd guideline. While it's certainly convenient to launch
> a program using commands like 'system' in perl, it's mostly better to
> use a library (or service). In fact the authors agree with this in 5,
> in which case why are we recommending them do this here?
> 
> The advice seems a bit confused in general. It's not really clear if
> the authors are saying if it's a good idea or bad idea to call external programs.
> 
> I think general advice to 'take care when calling external programs'
> may be more appropriate.

We've re-written this section more mindful of the package manager rule to follow and more explicitly pushing back against using external executables and services.

> built utility / package manager
> 
> Excellent advice that is all too often ignored.
> 
> As noted above the presentation in combo with 4 is
> confusing. Generally a package manager is used when an external
> library, module or package is re-used. But as noted there is generally
> no advice on reusing these artefacts, as 4 is about running programs.

We believe we have addressed the discrepencies between these two rules with our re-write.

> Do not require root privileges
> 
> I certainly wouldn't contradict this but I wonder why it's called
> out. There are many possible inadvisable things we've all seen in years of programming, but maybe this could be combined with advice about running external programs (4) or making the software runnable outside a default user environment (7)

(Greg, you have something for this? My motivation is painful experience.)
..................

> Eliminate hard-coded paths
> 
> Excellent advice, and worth calling out as it is such a frequent problem

:)

> Configuration from command line
> 
> Good advice, but could be combined with 7, as they are both about configuration.

................

> Include a small test set
> 
> Good advice, but there is no advice about including software unit
> tests. This is good general software engineering advice, and the
> existence of simple cloud-based CI frameworks like Travis that
> integrate with GitHub make it very easy to ensure that every change to
> a program does not break thing.
> 
> Produce identical results when given identical inputs
> 
> Generally good advice, but perhaps overly specific. There is a lengthy
> discussion about what to do in the case of algorithms that involve
> random numbers. What about software that calls external services?
> 
> In my opinion this is a judgment call by the software
> author. Certainly, if there is no deliberate randomization or calling
> of external services then nondetermistic output would be troubling.

If you call external services, then according to the (new!) guideline in "Reuse software", that software should itself be robust and therefore produce the same results on every run. Some services are updated to change, add or remove data or results and this is unfortunate, but expected. Two subsequent runs of the software, barring any other changes, should still produce identical results.

In addition, software that doesn't produce identical results is much more difficult to test. Depending on the amount of variation, it can be nearly impossible.

..................
(Greg, any other changes you want to make here?)

> What we left out
> 
> As mentioned above, documentation is important, and some of the
> existing 10 could be merged to make room for this.

We merged all mentions of documentation into the first rule.

> For services, I am not sure why authentication is mentioned, even in
> passing. It doesn't really pertain to robustness per se.

This reference has been removed.

> Reviewer 3
> 
> Comments to the author:
> 
> This paper sets out to describe 10 simple rules which, if applied, can contribute to making software research more robust. It provides a useful set of guidelines that should generally improve the robustness of research software and it should be of interest to most graduate students who are writing software in the process of their research.
> 
> I recommend the publication of this paper but would request the addition or additional emphasis on one point which is not called out as a rule but which is mentioned in passing in several places within the paper. In particular I am referring to the recommendation to put all software into a source control system. As this is an opinion paper, I do not think this is a necessary change for publication (in my opinion it’s a vital rule, that is just my opinion) but it is referenced in several places and in particular the author’s definition of “robust software” includes “...is kept under version control”. As this is an issue frequently ignored or addressed “too late” by graduate students I would like to see this get additional emphasis, even if it is not promoted to being a full rule.

..............

> The remainder of my comments are meant to provide (hopefully) useful feedback to the authors in the line of suggestions that may enhance the manuscript.
> 
> In describing successful software research projects the authors state “...who have adopted strong software engineering approaches, have high standards of reproducibility, use good testing practices, and foster inter-institutional collaborations that grow into strong user bases”.. All of these points can be clearly tied to their rules with the exception of the final statement about inter-institutional collaborations. While such collaborations are desirable, particularly to establish project continuity and longevity, they are not strictly necessary for robustness and this is not further referenced in the manuscript.

We removed this line from the text.

> In several cases, it seems that the paper might violate the first rule of “The 10 simple rules for writing a 10 simple rules paper” (http://journals.plos.org/ploscompbiol/article?id=10.1371/journal.pcbi.1003858) with multiple rules combined into a single rule.

We reviewed all the sections you mentioned. In general, we believe that most of our *bolded* sections within each rule are refinements of the rule and not separate rules. We address each one of your concerns in-line below.

> For example rule 8 combines putting parameters on the command line with echoing all parameters to the log file to support reproducibility and provenance.

................

> Similarly rule 5 covers both automating the build process and automating the deployment and installation.

..............

> Rule 4 also tends to stuff multiple rules (bolded) into one rule (( limit dependencies, ensure dependencies are available) these 2 are closely related), with “ use native functions” which is a very different idea and not what I would expect from a ‘reuse’ rule.

We removed this last 'native functions' rule and replaced it with something more appropriate.

> Finally the ‘What we left out’ section implicitly defines an additional rule (documentation). In this particular case there could have been a single rule (Document your code and usage) that could encompass the points of both Rules 1 and 2 (Readme, usage) as well as the points from ‘What we left out’.

We combined all mentions of documentation into a single rule.

> In rule 8, I would tend to suggest generalizing the rule somewhat to suggest externalizing parameters rather than necessarily requiring them to be available as command line options. For example if a configuration file is used to provide all parameter values, this could accomplish the task of separating the parameters out from the code as well as providing a record of the values used for a run as they describe through echoing parameters to a log file.

....

### 2017-01-13 robust-software.tex
Updating rule numbers and references, and making space for a new rule about version control.
1.  Added a new "Rule 1: Use version control" as a placeholder.
2.  Updated rule numbers for all other rules.
3.  Checked cross-rule references.
4.  Added new commands `\rulemajor` and `\ruleminor` to make rules and sub-rules easier to find.
5.  Updated the checklist to match the new rule set.
6.  Removed the "what we left out" section - it now had only one entry, so it looked a little lonely, and the full list of what we've left out would be very long.
7.  Removed the "bonus rules" section of the chec

%% Include all macros below
\newcommand{\fixme}[2]{\textsc{\textbf{{#1}: {#2}}}}
\newcommand{\recommend}[1]{\textit{#1}}
\newcommand{\withurl}[2]{{#1}\footnote{\texttt{#2}}}

@@

Note: practices \emph{not} included in the list below include version control,
code review, unit testing, and other practices essential to maintainability and
sustainable software development.
These elements are essential to building software of any kind
\cite{wilson2014,wilson2016}.
However,
our focus in this short guide is on the software that is already built,
not on the process used to build it.

\section{Document your code and usage}

@@

\item[\textbf{Explain what the software does.}] 
There's nothing more frustrating
than downloading and installing something only to
find out that it doesn't do what you thought it did.

\item[\textbf{List required dependencies.}] Often, software depends on
specific versions of libraries, modules, or operating systems. Include the name and
version number for each dependency or a link to the requirements file. 
We address dependencies in more detail in rule 5.

\item[\textbf{Provide compilation/installation instructions:}]
If the software needs to be compiled
or installed, list those instructions in the README.

\item[\textbf{List input and output files.}] All possible input and output files
should be listed in this section. 
If they use a standard format, link to the specification and
version. If you extend the standard or have your own format, define it
explicitly, listing all the required fields and acceptable values.
If there is no rigorous format, as is common with log files, show at least a
few lines from an
example file and explain what the sections mean.

@@

\item[\textbf{State attributions and licensing.}] Attributions are how you credit
your main contributors; licenses are how others may use and
credit your work.
Leave no
question in anyone's mind about whether your software can be used
commercially, how much modification is permitted, and how other software
needs to credit you. If your software is not open source, state that clearly.
\end{description}

In addition to a README, a robust program should
\textbf{print usage information when launching from the command line that explains the software's features}.
Usage information provides the first line of help for both first-time and
experienced users of command-line applications. Terseness is
important: usage that extends for multiple screens is 
difficult to read or refer to on the fly.

@@

\item[\textbf{The syntax for running the program}] This includes the
  name of the program and defines the relative location of optional
  and required flags, arguments and values for execution.  Arguments
  in {[}square brackets{]} are usually optional. An ellipsis (\ldots
  e.g. ``{[}OPTION{]}\ldots{}'') indicates that more than one value
  can be provided.

\item[\textbf{Description}] Similar to the README, the description
  reminds users of the software's primary function.

\item[\textbf{Most commonly used arguments, a description of each, and
    the default values}] Not all arguments need to appear in the
  usage, but those most commonly used should be listed. Users will
  rely on this for quick reference when working with the software.

\item[\textbf{Where to find more information}] Whether it's an email
  address, web site, or manual page, there should be an indication
  where the user can go to find out more.

\item[\textbf{Printed to standard output}] Usage should be displayed
  on standard output so that it can be piped into \texttt{less},
  searched with \texttt{grep}, or compared to the previous version
  with \texttt{diff}.

\item[\textbf{Exit with an appropriate exit code}] When usage is
  invoked by providing incorrect arguments, the program should exit
  with a non-zero code to indicate an error. However, when help is
  explicitly requested, the software should not exit with an error.

@@

\section{Make common operations easy to control.}

Being able to change parameters on the fly to determine if and how
they change the results is important as your software gains more users,
as it facilitates exploratory analysis and parameter sweeping.
Programs should therefore
\textbf{allow the most commonly changed parameters to be configured from the command line}.

@@

\textbf{Check that all input values are in a reasonable range at startup}. 
Few things are as annoying as having a program announce after running for two
hours that it isn't going to save its results because the requested directory
doesn't exist.

To make programs even easier to use,
\textbf{choose reasonable defaults where they exist}
and \textbf{set no defaults at all when there aren't any reasonable ones}.
You can set reasonable default values
as long as any command line arguments
override those values.

@@

\section{Version your releases.}

@@

Most software has a version number composed of a decimal number that
increments as new versions are released. 
There are many different ways
to construct and interpret this number, but most importantly for us, a
particular software version run with the same parameters should give
identical results no matter when it's run. Results include both correct
output as well as any errors.
\textbf{Increment your version number every time you release your software to
other people}.

@@

\textbf{The version of your software should be easily available by 
supplying \texttt{-\/-version} or \texttt{-v} on the command line}. This command should print
the software name and version number, and it should
also be \textbf{included in all of the program's output}, particularly debugging
traces.  If someone needs help, it's important that they be able to tell
whoever's helping them which version of the software they're using.

While new releases may make a program better in general,
they can simultaneously create work for someone
who integrated the old version into their own workflow a year or two ago,
and won't see any benefits from upgrading.
A program's authors should therefore \textbf{ensure that old released versions
continue to be available.}
A number of mechanisms exist for
controlled release that range from as simple as adding an appropriate
commit message or tag to version control, to official releases alongside
code on Bitbucket or GitHub, to depositing into a
repository like apt, yum, homebrew, CPAN, etc. Choose the method that
best suits the number and expertise of users you anticipate.

\section{Reuse software (within reason).}

@@

First, 
\textbf{make sure that you really need the auxiliary program}. If you are
executing GNU sort instead of figuring out how to sort lists in Python,
it may not be worth the pain of integration. Reuse software that offers some
measurable improvement to your project.

Second, if launching an executable, \textbf{ensure the appropriate software and version is available}.
Either allow the user to
configure the exact path to the package, distribute the program with the
dependent software, or download it during installation using your
package manager. If the executable requires internet access, check for that
early in execution.

Third, \textbf{ensure that reused software is robust}. Relying on erratic third
party libraries or software is a recipe for tears. Prefer software that follows
good software development practices, is open for support questions, and is
available from a stable location or repository using your package manager.

@@

\section{Rely on build tools and package managers for installation.}

To compile code, deploy applications, and automate other tasks, 
programmers routinely use build tools like Make, Rake, Maven, Ant or MS Build.
These tools can also be used to manage runtime environments,
i.e.,
to check that the right versions of required packages are installed
and install or upgrade them if they are not.
As mentioned in the previous rule,
a package manager can mitigate some of the difficulties in software reuse.

The same tools can and should be used to manage runtime environments on users' machines as well.
Accordingly,
developers should
\textbf{document all dependencies in a machine-readable form}.
Package managers like apt and yum are available on most Unix-like systems, and
application package managers exist for specific languages like Python (pip),
Java (Maven/Gradle), and Ruby (RubyGems). These package managers can be used
together with the build utility to ensure that dependencies are available at
compile/run time.

@@

Conversely, developers should
\textbf{avoid depending on scripts and tools which are not available as packages}.
In many cases, a program's author may not realize that some tool was built locally, and
doesn't exist elsewhere. At present, the only sure way to discover such
unknown dependencies is to install on a system administered by someone
else and see what breaks. As use of virtualization containers becomes more
widespread, software installation can also be tested on a virtual machine or
container system like Docker.

\section{Do not require root or other special privileges.}

@@

Installing or running a program with root privileges is often
convenient, since doing so automatically bypasses all those pesky safety
checks that might otherwise get in the user's way. However, those checks
are there for a reason: scientific software packages may not
intentionally be malware, but one small bug or over-eager file-matching
expression can certainly make them behave as if they were. Outside of
very unusual circumstances,
\textbf{packages should not require root privileges to set up or use}.

@@

Developers should therefore
\textbf{allow packages to be installed in an arbitrary location},
e.g., under a user's home directory in
\texttt{\textasciitilde{}/packagename}, or in directories with standard
names like \texttt{bin}, \texttt{lib}, and \texttt{man} under a chosen
directory. If the first option is chosen, the user may need to modify
her search path to include the package's executables and libraries, but
this can (more or less) be automated, and is much less risky than
setting things up as root.

Testing the ability to install software has traditionally been regarded as difficult,
since it necessarily alters the machine on which the test is conducted.
Lightweight virtualization containers like Docker make this much easier as well, 
or simply \textbf{ask another person to try and build your software}.

\section{Eliminate hard-coded paths.}

@@

For these reasons, users should be able to
\textbf{set the names and locations of input and output files as command-line parameters}.
This rule applies to reference data sets as well as the user's own
data: if a user wants to try a new gene identification algorithm using
a different set of genes as a training set, she should not have to
edit the software to do so.
A corollary to this rule is
\textbf{do not require users to navigate to a particular directory to do their work},
since ``where I have to be'' is just another hard-coded path.

@@

\section{Include a small test set that can be run to ensure the software is actually working.}

@@

In order to be useful, \textbf{make the test script easy to find and run}. 
Many build systems will also run unit tests if provided them at compile time.
If the build system is not amenable to testing, a 
single file in the project's root directory named \texttt{runtests.sh}
or something equally obvious is a much better solution than documenting
test cases and requiring people to copy and paste them.

Equally, \textbf{make the test script's output easy to interpret}. Screens
full of correlation coefficients do not qualify: instead, the script's
output should be simple to understand for non-experts,
such as one line per test, with the test's name
and its pass/fail status, followed by a single summary line saying how
many tests were run and how many passed or failed. If many or all tests
fail because of missing dependencies, that fact should be displayed
once, clearly, rather than once per test, so that users have a clear
idea of what they need to fix and how much work it's likely to take.

@@

\section{Produce identical results when given identical inputs.}

The usage message tells users what the program could do.
It is equally important for the program to tell users what it actually did.
Accordingly,
when the program starts, it should \textbf{echo all parameters and software
versions to standard out or a log file alongside the results}. This
feature supports greater reproducibility because any result can be
replicated with only the previous output files as reference.

Given a set of parameters and a dataset, \textbf{a particular version of a program
should produce the same results every time it is run}
to aid testing, debugging, and reproducibility.
Even minor changes to code can cause minor changes in output because of floating-point issues,
which means that getting exactly the same output for the same input and parameters
probably won't work during development,
but it should still be a goal for people who have deployed a specific version.

@@

Given the size of biological data, it is unreasonable to suggest that
random algorithms be removed. However, most programs use a pseudo-random
number generator, which uses a starting seed and an equation to
approximate random numbers. Setting the seed to a consistent value
can remove randomness between runs. \textbf{Allow the user to optionally provide
the random seed as an input parameter}, thus rendering the program deterministic
for those cases where it matters. If the seed is set internally (e.g.,
using clock time), echo it to the output for re-use later. 
If setting the seed is not possible, \textbf{make sure the acceptable tolerance is
known and detailed in documentation and in the tests}.

\section*{What We Left Out}

Any short list of rules for developing robust software will
necessarily be incomplete.   Our more significant omissions
include:

\item[\textbf{Services}] The rules we present above are necessary, but
  not sufficient, for software that needs to interact with services
  such as databases and web servers, or coordinates multiple hosts or interfaces, 
  and is out of scope of this paper.

>>>>

%% Include all macros below
\newcommand{\fixme}[2]{\textsc{\textbf{{#1}: {#2}}}}
\newcommand{\recommend}[1]{\textit{#1}}
\newcommand{\withurl}[2]{{#1}\footnote{\texttt{#2}}}
\newcommand{\rulemajor}[1]{\section{#1}}
\newcommand{\ruleminor}[1]{\textbf{#1}}

@@

\rulemajor{Rule 1: Use version control.}

FIXME: placeholder while renumbering.

Our focus in this short guide is on the software that is already built,
not on the process used to build it.
However,
we believe that version control is essential to sustainable software development
\cite{wilson2014,wilson2016}.

\rulemajor{Rule 2: Document your code and usage}

@@

\item[\ruleminor{Explain what the software does.}] 
There's nothing more frustrating
than downloading and installing something only to
find out that it doesn't do what you thought it did.

\item[\ruleminor{List required dependencies.}] Often, software depends on
specific versions of libraries, modules, or operating systems. Include the name and
version number for each dependency or a link to the requirements file. 
We address dependencies in more detail in Rule~5.

\item[\ruleminor{Provide compilation/installation instructions:}]
If the software needs to be compiled
or installed, list those instructions in the README.

\item[\ruleminor{List input and output files.}] All possible input and output files
should be listed in this section. 
If they use a standard format, link to the specification and
version. If you extend the standard or have your own format, define it
explicitly, listing all the required fields and acceptable values.
If there is no rigorous format, as is common with log files, show at least a
few lines from an
example file and explain what the sections mean.

@@

\item[\ruleminor{State attributions and licensing.}] Attributions are how you credit
your main contributors; licenses are how others may use and
credit your work.
Leave no
question in anyone's mind about whether your software can be used
commercially, how much modification is permitted, and how other software
needs to credit you. If your software is not open source, state that clearly.
\end{description}

In addition to a README, a robust program should
\ruleminor{print usage information when launching from the command line that explains the software's features}.
Usage information provides the first line of help for both first-time and
experienced users of command-line applications. Terseness is
important: usage that extends for multiple screens is 
difficult to read or refer to on the fly.

@@

\item[\ruleminor{The syntax for running the program}] This includes the
  name of the program and defines the relative location of optional
  and required flags, arguments and values for execution.  Arguments
  in {[}square brackets{]} are usually optional. An ellipsis (\ldots
  e.g. ``{[}OPTION{]}\ldots{}'') indicates that more than one value
  can be provided.

\item[\ruleminor{Primary function}] Similar to the README, the description
  reminds users of the software's primary function.

\item[\ruleminor{Most commonly used arguments, a description of each, and
    the default values}] Not all arguments need to appear in the
  usage, but those most commonly used should be listed. Users will
  rely on this for quick reference when working with the software.

\item[\ruleminor{Where to find more information}] Whether it's an email
  address, web site, or manual page, there should be an indication
  where the user can go to find out more.

\item[\ruleminor{Printed to standard output}] Usage should be displayed
  on standard output so that it can be piped into \texttt{less},
  searched with \texttt{grep}, or compared to the previous version
  with \texttt{diff}.

\item[\ruleminor{Exit with an appropriate exit code}] When usage is
  invoked by providing incorrect arguments, the program should exit
  with a non-zero code to indicate an error. However, when help is
  explicitly requested, the software should not exit with an error.

@@

\rulemajor{Rule 3: Make common operations easy to control.}

Being able to change parameters on the fly to determine if and how
they change the results is important as your software gains more users,
as it facilitates exploratory analysis and parameter sweeping.
Programs should therefore
\ruleminor{allow the most commonly changed parameters to be configured from the command line}.

@@

\ruleminor{Check that all input values are in a reasonable range at startup}. 
Few things are as annoying as having a program announce after running for two
hours that it isn't going to save its results because the requested directory
doesn't exist.

To make programs even easier to use,
\ruleminor{choose reasonable defaults where they exist}
and \ruleminor{set no defaults at all when there aren't any reasonable ones}.
You can set reasonable default values
as long as any command line arguments
override those values.

@@

\rulemajor{Rule 4: Version your releases.}

@@

Most software has a version number composed of a decimal number that
increments as new versions are released. 
There are many different ways
to construct and interpret this number, but most importantly for us, a
particular software version run with the same parameters should give
identical results no matter when it's run. Results include both correct
output as well as any errors.
\ruleminor{Increment your version number every time you release your software to
other people}.

@@

\ruleminor{The version of your software should be easily available by 
supplying \texttt{-\/-version} or \texttt{-v} on the command line}. This command should print
the software name and version number, and it should
also be \ruleminor{included in all of the program's output}, particularly debugging
traces.  If someone needs help, it's important that they be able to tell
whoever's helping them which version of the software they're using.

While new releases may make a program better in general,
they can simultaneously create work for someone
who integrated the old version into their own workflow a year or two ago,
and won't see any benefits from upgrading.
A program's authors should therefore \ruleminor{ensure that old released versions
continue to be available.}
A number of mechanisms exist for
controlled release that range from as simple as adding an appropriate
commit message or tag to version control~\cite{blischak2016}, to official releases alongside
code on Bitbucket or GitHub, to depositing into a
repository like apt, yum, homebrew, CPAN, etc. Choose the method that
best suits the number and expertise of users you anticipate.

\rulemajor{Rule 5: Reuse software (within reason).}

@@

First, 
\ruleminor{make sure that you really need the auxiliary program}. If you are
executing GNU sort instead of figuring out how to sort lists in Python,
it may not be worth the pain of integration. Reuse software that offers some
measurable improvement to your project.

Second, if launching an executable, \ruleminor{ensure the appropriate software and version is available}.
Either allow the user to
configure the exact path to the package, distribute the program with the
dependent software, or download it during installation using your
package manager. If the executable requires internet access, check for that
early in execution.

Third, \ruleminor{ensure that reused software is robust}. Relying on erratic third
party libraries or software is a recipe for tears. Prefer software that follows
good software development practices, is open for support questions, and is
available from a stable location or repository using your package manager.

@@

\rulemajor{Rule 6: Rely on build tools and package managers for installation.}

To compile code, deploy applications, and automate other tasks, 
programmers routinely use build tools like Make, Rake, Maven, Ant or MS Build.
These tools can also be used to manage runtime environments,
i.e.,
to check that the right versions of required packages are installed
and install or upgrade them if they are not.
As mentioned in Rule~5,
a package manager can mitigate some of the difficulties in software reuse.

The same tools can and should be used to manage runtime environments on users' machines as well.
Accordingly,
developers should
\ruleminor{document all dependencies in a machine-readable form}.
Package managers like apt and yum are available on most Unix-like systems, and
application package managers exist for specific languages like Python (pip),
Java (Maven/Gradle), and Ruby (RubyGems). These package managers can be used
together with the build utility to ensure that dependencies are available at
compile/run time.

@@

Conversely, developers should
\ruleminor{avoid depending on scripts and tools which are not available as packages}.
In many cases, a program's author may not realize that some tool was built locally, and
doesn't exist elsewhere. At present, the only sure way to discover such
unknown dependencies is to install on a system administered by someone
else and see what breaks. As use of virtualization containers becomes more
widespread, software installation can also be tested on a virtual machine or
container system like Docker.

\rulemajor{Rule 7: Do not require root or other special privileges to install or run.}

@@

Installing or running a program with root privileges is often
convenient, since doing so automatically bypasses all those pesky safety
checks that might otherwise get in the user's way. However, those checks
are there for a reason: scientific software packages may not
intentionally be malware, but one small bug or over-eager file-matching
expression can certainly make them behave as if they were. Outside of
very unusual circumstances,
\ruleminor{packages should not require root privileges to set up or use}.

@@

Developers should therefore
\ruleminor{allow packages to be installed in an arbitrary location},
e.g., under a user's home directory in
\texttt{\textasciitilde{}/packagename}, or in directories with standard
names like \texttt{bin}, \texttt{lib}, and \texttt{man} under a chosen
directory. If the first option is chosen, the user may need to modify
her search path to include the package's executables and libraries, but
this can (more or less) be automated, and is much less risky than
setting things up as root.

Testing the ability to install software has traditionally been regarded as difficult,
since it necessarily alters the machine on which the test is conducted.
Lightweight virtualization containers like Docker make this much easier as well, 
or simply \ruleminor{ask another person to try and build your software before releasing it}.

\rulemajor{Rule 8: Eliminate hard-coded paths.}

@@

For these reasons, users should be able to
\ruleminor{set the names and locations of input and output files as command-line parameters}.
This rule applies to reference data sets as well as the user's own
data: if a user wants to try a new gene identification algorithm using
a different set of genes as a training set, she should not have to
edit the software to do so.
A corollary to this rule is
\ruleminor{do not require users to navigate to a particular directory to do their work},
since ``where I have to be'' is just another hard-coded path.

@@

\rulemajor{Rule 9: Include a small test set that can be run to ensure the software is actually working.}

@@

In order to be useful, \ruleminor{make the test script easy to find and run}. 
Many build systems will also run unit tests if provided them at compile time.
If the build system is not amenable to testing, a 
single file in the project's root directory named \texttt{runtests.sh}
or something equally obvious is a much better solution than documenting
test cases and requiring people to copy and paste them.

Equally, \ruleminor{make the test script's output easy to interpret}. Screens
full of correlation coefficients do not qualify: instead, the script's
output should be simple to understand for non-experts,
such as one line per test, with the test's name
and its pass/fail status, followed by a single summary line saying how
many tests were run and how many passed or failed. If many or all tests
fail because of missing dependencies, that fact should be displayed
once, clearly, rather than once per test, so that users have a clear
idea of what they need to fix and how much work it's likely to take.

@@

\rulemajor{Rule 10: Produce identical results when given identical inputs.}

The usage message tells users what the program could do.
It is equally important for the program to tell users what it actually did.
Accordingly,
when the program starts, it should \ruleminor{echo all parameters and software
versions to standard out or a log file alongside the results}. This
feature supports greater reproducibility because any result can be
replicated with only the previous output files as reference.

Given a set of parameters and a dataset, \ruleminor{a particular version of a program
should produce the same results every time it is run}
to aid testing, debugging, and reproducibility.
Even minor changes to code can cause minor changes in output because of floating-point issues,
which means that getting exactly the same output for the same input and parameters
probably won't work during development,
but it should still be a goal for people who have deployed a specific version.

@@

Given the size of biological data, it is unreasonable to suggest that
random algorithms be removed. However, most programs use a pseudo-random
number generator, which uses a starting seed and an equation to
approximate random numbers. Setting the seed to a consistent value
can remove randomness between runs. \ruleminor{Allow the user to optionally provide
the random seed as an input parameter}, thus rendering the program deterministic
for those cases where it matters. If the seed is set internally (e.g.,
using clock time), echo it to the output for re-use later. 
If setting the seed is not possible, \ruleminor{make sure the acceptable tolerance is
known and detailed in documentation and in the tests}.
